Measurement Bias from Address Aliasing

### 2014-08-26 paper.tex
Updated the setup table.

Because bias can be hardware dependent, and to keep the scope of this work manageable, we initially chose to focus our analysis on the 4th generation Intel ``Haswell'' microarchitecture specifically.
We later reproduced similar results on the previous ``Sandy Bridge'' CPU, as well as an older Core2 Duo machine.
Each setup is described in Table~\ref{tab:machines}, and referenced throughout the paper.

\begin{table}
  \tbl{Experimental setups used\label{tab:machines}}{
  \begin{tabular}{|l|l|}
    \hline
Core          & Intel\textsuperscript{\textregistered{}} Core\texttrademark{}2 Duo P8600 (laptop) \\
              & 64-bit Ubuntu 14.04 LTS (Linux 3.13.0), GCC 4.8.2-19ubuntu1 \\ \hline
Sandy Bridge  & Intel\textsuperscript{\textregistered{}} Core\texttrademark{} i7-3???? \\
              & Ubuntu ?? \\ \hline
Haswell       & Intel\textsuperscript{\textregistered{}} Core\texttrademark{} i7-4770K \\
              & 64-bit Ubuntu 14.04 LTS (Linux 3.13.0), GCC 4.8.2-19ubuntu1 \\ \hline
  \end{tabular}}
\end{table}

We also have to ensure we are not affected by measurement bias beyond what we are actually trying to observe, and in general follow best practices on gathering data~\cite{Mytkowicz:2009:WrongData}.
Most importantly, this means keeping the memory address space under control.
For security reasons, addresses of stack, heap and dynamic libraries are often randomized at load time, a technique known as \emph{Address Space Layout Randomization}~\cite{Pax:ASLR,Bhatkar:AddressObfuscation}. 
By disabling ASLR, we are able to execute the same program multiple times with identical virtual address spaces.
All experiments are done on a dedicated machine under minimal load, with \emph{frequency scaling} disabled to keep the CPU's clock speed fixed.
Finally, we disable \emph{Hyper-threading} (HT), which reduces the possibility for resource contention between threads.
Some PMU events also require HT to be turned off, or produce inaccurate results with HT enabled.

>>>>

Because bias can be hardware dependent, and to keep the scope of this work manageable, we initially chose to focus our analysis on the 4th generation Intel ``Haswell'' microarchitecture specifically.
We later reproduced similar results on the previous ``Nehalem'', ``Sandy Bridge'' and ``Ivy Bridge'' CPUs, as well as an older Core2 Duo machine.
Each setup is described in Table~\ref{tab:machines}, and referenced throughout the paper. The compiler used is GCC 4.8.2-19ubuntu1 and the same binary is tested on all systems.

\begin{table}
  \tbl{Experimental setups used\label{tab:machines}}{
  \begin{tabular}{|l|l|}
    \hline
Core          & Intel\textsuperscript{\textregistered{}} Core\texttrademark{}2 Duo P8600 (laptop) \\
              & 64-bit Ubuntu 14.04 LTS (Linux 3.13.0) \\ \hline
Nehalem       & Intel\textsuperscript{\textregistered{}} Core\texttrademark{} i7-950 \\
              & 64-bit Ubuntu 12.04 LTS (Linux 3.2.0) \\ \hline
Ivy Bridge    & Intel\textsuperscript{\textregistered{}} Core\texttrademark{} i5-3470 \\
              & 64-bit Ubuntu 12.04 LTS (Linux 3.8.0) \\ \hline
Sandy Bridge  & Intel\textsuperscript{\textregistered{}} Core\texttrademark{} i7-2670QM (laptop) \\
              & 64-bit Ubuntu 11.04 (Linux 3.0.0) \\ \hline
Haswell       & Intel\textsuperscript{\textregistered{}} Core\texttrademark{} i7-4770K \\
              & 64-bit Ubuntu 14.04 LTS (Linux 3.13.0) \\ \hline
  \end{tabular}}
\end{table}

We also have to ensure we are not affected by measurement bias beyond what we are actually trying to observe, and in general follow best practices on gathering data~\cite{Mytkowicz:2009:WrongData}.
Most importantly, this means keeping the memory address space under control.
For security reasons, addresses of the stack, heap and dynamic libraries are often randomized at load time, a technique known as \emph{Address Space Layout Randomization}~\cite{Pax:ASLR,Bhatkar:AddressObfuscation}. 
By disabling ASLR, we are able to execute the same program multiple times with identical virtual address spaces.
All experiments are done on a dedicated machine under minimal load, with \emph{frequency scaling} disabled to keep the CPU's clock speed fixed.
Finally, we disable \emph{Hyper-threading} (HT), which reduces the possibility for resource contention between threads.
Some PMU events also require HT to be turned off, or produce inaccurate results with HT enabled.
